What gives better mileage? Automatic or Manual transmission
========================================================

# Executive summary
    
In this report, we examine the dataset from 1974 Motor Trend US magazine to determine the effect of transmission type on miles per gallon (mpg). We conducted a t-test for the difference between mean mpg of automatic and manual transmission cars. At 95% confidence interval, it is observed that there is a difference in mean of about 7 miles more for the manual transmitted cars. We then quantified the difference in mpg between automatic and manula transmissions using a liner regression model. This model shows that a manual transmission increase the expected mpg by about 2.9 miles when all the other factors are kept constant.

## Cleaning data

Load and look at data

```{r}
data(mtcars)
str(mtcars)
```

Convert "cyl", "vs", "am", gear", "carb" variables into factors.

```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
```

Add readable levels to variable "am".

```{r}
levels(mtcars$am) <- c("Auto", "Manual")
```


```{r}
summary(cars)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

